---
title: "Analysis-Predit Churning Bank Customers"
author: "Anjana Molugu"
date: "03/01/2021"
output:
   html_document:
    code_folding: hide
    highlight: tango
    toc: yes
    toc_float:
      collapsed: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 

```{r,echo=F, results='hide', message=F, warning=F}
library(dplyr,warn.conflicts=F, quietly=T)
library(plotly,warn.conflicts=F, quietly=T)
library(ggplot2,warn.conflicts=F, quietly=T)
library(gridExtra,warn.conflicts=F, quietly=T)
library(plotrix,warn.conflicts=F, quietly=T)
library(tidyverse,warn.conflicts=F, quietly=T)
library(sampling,warn.conflicts=F, quietly=T)
```

 

```{r echo=F, results='hide', message=F, warning=F}
creditData <- read.csv("BankChurners.csv" )

## Removing two columns which are output of a Classifier

creditData = creditData %>% 
                select(-c(Naive_Bayes_Classifier_Attrition_Flag_Card_Category_Contacts_Count_12_mon_Dependent_count_Education_Level_Months_Inactive_12_mon_1
                          ,Naive_Bayes_Classifier_Attrition_Flag_Card_Category_Contacts_Count_12_mon_Dependent_count_Education_Level_Months_Inactive_12_mon_2))

## Listing all the Categorical Variables and removing few categories 
unique(creditData$Marital_Status)
### filtering Marital Status 'Unknown'
creditData=creditData[creditData$Marital_Status!='Unknown',]
```


```{r}

creditData <-creditData %>% drop_na()

 data(creditData)
 names(creditData)
names(sort(table(creditData$Marital_Status), decreasing = TRUE  ) [1:3]) ->martialStatusTop5

martialStatusData <- subset(creditData, Marital_Status %in% martialStatusTop5)

set.seed(1206)



n <- 50
N <- nrow(martialStatusData) 
```


### Simple random sampling without replacement for 10 samples


```{r}
s <- srswor(50, nrow(creditData))
sample.1 <- creditData[s != 0, ]
table(sample.1$Credit_Limit)

 srswor1 <- plot_ly(as.data.frame(sample.1$Credit_Limit ), x = ~sample.1$Credit_Limit) %>%
  add_histogram(name = "srswor") 
 mean(sample.1$Credit_Limit)

 
 
 
```


### Systematic sampling

```{r}
k <- ceiling(N / n)


r <- sample(k, 1)

# select every kth item
s <- seq(r, by = k, length = n)
sample.2 <- creditData[s, ]



 sys <- plot_ly(as.data.frame(sample.2$Credit_Limit), x = ~sample.2$Credit_Limit) %>%
  add_histogram(name = "Systematic Sampling") 
mean(sample.2$Credit_Limit)

```

### Inclusion Probabilities

For unequal probabilities, the inclusionprobabilities function computes the probability for each item to be included in the sample with probabilities proportional to the size.

```{r}

pik <- inclusionprobabilities(martialStatusData$Credit_Limit,50)

sumPik3 <- sum(pik)
s <- UPsystematic(pik)
sample.3 <- martialStatusData[s!=0,]
table3 <-table(sample.3$Credit_Limit)
#Percentage of respective Department
Percentage <- (table(sample.3$Credit_Limit)/50)*100
 ip <- plot_ly(as.data.frame(sample.3$Credit_Limit), x = ~sample.3$Credit_Limit) %>%
  add_histogram(name = "Inclusion Probabilities") 
 mean(sample.3$Credit_Limit)
 
```

### Stratified sampling

In stratifies sampling, the items from the frame are subdivided into separate N subgroups based on some common characteristic, e.g., gender, race, year of school, etc. The subgroups are known as strata

```{r}
order.index <- order(martialStatusData$Marital_Status)
data <- martialStatusData[order.index, ]  

freq <- table( martialStatusData$Marital_Status)
sizes <- round(50 * freq / sum(freq))

sum(sizes)
st_Martial_status <-  strata(data, stratanames = c("Marital_Status"),
                         size = sizes, method = "srswor")

sample.4 <-getdata(data,st_Martial_status)
table4 <-table(sample.4$Education_Level)
#Percentage of respective Department
Percentage <- (table(sample.4$Education_Level)/50)*100

 ss <- plot_ly(as.data.frame(sample.4$Credit_Limit), x = ~sample.4$Credit_Limit) %>%
  add_histogram(name = "Stratified sampling") 
 mean(sample.4$Credit_Limit)
 
 subplot(srswor1, sys, ip,  ss,
  nrows = 4, shareX = TRUE
  
)



```

```{r}
   cat("Population sample mean       : ", mean(creditData$Credit_Limit) ,
  "\nSimple random  mean    : ",    mean(sample.1$Credit_Limit),
  "\nSystematic sampling mean          : ", mean(sample.2$Credit_Limit),
  "\nInclusion Probabilities mean : ", mean(sample.3$Credit_Limit),
  "\nStratified sampling mean     : ", mean(sample.4$Credit_Limit))

```


## Conclusion
The dataset is imbalanced with 84:16 ratio of existing and attired customers
 
  •	There are more samples of females in dataset compared to males but the percentage of difference is not that significant so we can say that genders are uniformly distributed.
  
  •	If assuming that most of the customers with unknown education status lack any sort of education, we can state that more than 70% of the customers have a formal education level of         which about 35% have a higher level of education.
  
  •	Almost half of the customers of the bank is married and almost the entire other half are customers who are single. Only about 7% of the customers are divorced.



